Wide binaries in Gaia DR1

### 2016-09-24 paper/ms.tex
infer distances too

>>>>

\section{Methods} \label{sec:methods}

For a pair of stars $(i,j)$, we want to ratio the fully marginalized
likelihood (FML) of the stars having the same tangential velocity (numerator)
compared to the FML of the stars having different tangential velocities:
\begin{equation}
  \frac{p(\propm_i, \propm_j)_{\vperp_i = \vperp_j}}
  {p(\propm_i, \propm_j)_{\vperp_i \neq \vperp_j}}.
\end{equation}
When this ratio is large, the pair is a candidate binary star. These likelihood
expressions are marginalized over true tangential velocity and distance for each
star in the pair. For the numerator,
\begin{equation}
  p(\propm_i, \propm_j)_{\vperp_i = \vperp_j} =
    \int \, \dd \vperp \, \dd d_i \, \dd d_j \,
    p(\propm_i, \varpi_i, \vperp, d_i, \matrx{C}_i) \,
    p(\propm_j, \varpi_j, \vperp, d_j, \matrx{C}_j)
\end{equation}
where
\begin{align}
  p(\propm, \varpi, \vperp, d, \matrx{C}) &=
    p(\propm, \varpi \given \vperp, d, \matrx{C}) \, p(\vperp) \, p(d) \\
  p(\propm, \varpi \given \vperp, d, \matrx{C}) &=
    \left[\det\left(\frac{\matrx{C}^{-1}}{2\pi}\right)\right]^{1/2} \,
    \exp \left[ -\frac{1}{2} \transp{\left(\bs{x} - \bs{x}_\theta \right)} \,
    \matrx{C}^{-1} \,
    \left(\bs{x} - \bs{x}_\theta \right) \right] \\
  \bs{x} &= \transp{\left(\varpi, \mu_\alpha, \mu_\delta\right)} \\
  \bs{x}_\theta &= \transp{\left(\frac{1}{d}, \frac{v^\perp_\alpha}{d}, \frac{v^\perp_\delta}{d}\right)}
\end{align}
and we assume Gaussian priors on the tangential velocity components, $\vperp=(v^\perp_\alpha,v^\perp_\delta)$, and take the ``Milky Way distance prior'' from Coryn Bailer-Jones
\begin{align}
  p(v^\perp_\alpha,v^\perp_\alpha) = \mathcal{N}(0,25~\kms) \, \mathcal{N}(0,25~\kms)
\end{align}

For the denominator of the likelihood ratio,
\begin{equation}
  p(\propm_i, \propm_j)_{\vperp_i \neq \vperp_j} =
    \int \, \dd \vperp_i \, \dd \vperp_j \, \dd d_i \, \dd d_j \,
    p(\propm_i, \varpi_i, \vperp_i, d_i, \matrx{C}_i) \,
    p(\propm_j, \varpi_j, \vperp_j, d_j, \matrx{C}_j)
\end{equation}
